---
title: "Performance-Based HealthCare Provider Incentives"
author: "Asa Downs - SADA Systems"
date: "December 11, 2015"
output: 
  html_document:
    fig_width: 14
    fig_height: 10
    css: style.css
---


##Background

Exploratory Research was done on publicly available Medicare data as a proxy for more detailed Insurer and Provider data. The gaol was to explore the relationship between patient outcomes, provider efficiency, and insurer-provider compensation models.

##Exploratory Analysis

```{r getdata, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(knitr)
library(kfigr)

if (!file.exists('Hospital_Revised_Flatfiles.zip')) {
  download.file('https://data.medicare.gov/views/bg9k-emty/files/Ma46xU4I05xsIKuEqLLi-N-s7GoO2ZefzJ7SYyTIKjA?content_type=application%2Fzip%3B%20charset%3Dbinary&filename=Hospital_Revised_Flatfiles.zip','Hospital_Revised_Flatfiles.zip',method='curl')
  downloadDate <- date()  
}

if (!file.exists('raw')) {
  unzip('Hospital_Revised_Flatfiles.zip', exdir="./raw")
}

if (!exists('merged')) {
  readmissionsMortality <- tbl_df(read.csv('raw/Readmissions and Deaths - Hospital.csv', stringsAsFactors = FALSE, header = TRUE, na.strings=c("Not Available","Too Few To Report")))
  readmissionsMortality <- readmissionsMortality %>% select(one_of(c("Provider.ID","Measure.ID","Score"))) %>% 
    na.omit() %>% mutate(Provider.ID=as.integer(Provider.ID))
  
  readmissionHipKnee <- subset(readmissionsMortality, Measure.ID=='READM_30_HIP_KNEE')
  readmissionTotal <- subset(readmissionsMortality, Measure.ID=='READM_30_HOSP_WIDE')
  colnames(readmissionTotal) <- c("Provider.ID", "Measure.ID", "READM_SCORE")
  
  effectiveCare <- tbl_df(read.csv('raw/Timely and Effective Care - Hospital.csv', stringsAsFactors = FALSE, header = TRUE, na.strings=c("Not Available","Too Few To Report")))

  totalPerformanceScore <- tbl_df(read.csv('raw/hvbp_tps_08_06_2015.csv', stringsAsFactors = FALSE, header = TRUE, na.strings=c("Not Available","Too Few To Report")))
  totalPerformanceScore <- totalPerformanceScore %>% 
    select(one_of(c("Provider.Number","Weighted.Patient.Experience.of.Care.Domain.Score", "Weighted.Clinical.Process.of.Care.Domain.Score","Weighted.Outcome.Domain.Score","Weighted.Efficiency.Domain.Score","Hospital.Name","Total.Performance.Score","State"))) %>% 
    na.omit() %>% 
    mutate(Provider.Number=as.integer(Provider.Number)) %>% 
    group_by(State) %>% 
    filter(State=='FL') %>%
    gather(domain.type,domain.score, Weighted.Patient.Experience.of.Care.Domain.Score, Weighted.Clinical.Process.of.Care.Domain.Score, Weighted.Outcome.Domain.Score, Weighted.Efficiency.Domain.Score) %>%
    transform(Hospital.Name = reorder(Hospital.Name, domain.score))

  valueOfCare <- tbl_df(read.csv('raw/Payment and Value of Care - Hospital.csv', stringsAsFactors = FALSE, header = TRUE, na.strings=c("Not Available","Too Few To Report")))
  valueOfCare <- valueOfCare %>% select(one_of(c("Provider.ID","Value.of.care.display.name","Value.of.care.category","Payment.category"))) %>% 
    na.omit() %>% mutate(Provider.ID=as.integer(Provider.ID))
  
  spendingPerPatient <- tbl_df(read.csv('raw/Medicare Hospital Spending per Patient - Hospital.csv', stringsAsFactors = FALSE, header = TRUE, na.strings=c("Not Available","Too Few To Report")))
  spendingPerPatient <- spendingPerPatient %>% select(one_of(c("Provider.ID","Hospital.Name","City","State","ZIP.Code","County.Name","Score"))) %>% 
    na.omit()
  
  valueCareSpending <- inner_join(spendingPerPatient, valueOfCare, by="Provider.ID")
  
  readmissionSpendingScore <- inner_join(spendingPerPatient, readmissionTotal, by="Provider.ID")
}
```

We first wanted to confirm there was a relationship between readmission rates and Medicare hospital spending per patient. We fitted a simple linear model to these two pieces of data.

```{r readmissionScoreBeneficiaryScore, echo=TRUE}
readmissionModel <- lm(Score~READM_SCORE,data=readmissionSpendingScore)
summary(readmissionModel)
```

This data is plotted below.

```{r readmissionScoreBeneficiaryScoreGraph, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(readmissionSpendingScore, aes(x=READM_SCORE, y=Score)) + 
  geom_point() + 
  geom_smooth(method="lm")
```

We next wanted to explore how Medicare links payment to quality: 
> The HVBP program is part of CMS’ long-standing effort to link Medicare’s payment system to quality. The program implements value-based purchasing to the payment system that accounts for the largest share of Medicare spending, affecting payment for inpatient stays in over 3,500 hospitals across the country. Hospitals are paid for inpatient acute care services based on the quality of care, not just quantity of the services they provide.
>The adjusts hospitals’ payments based on their performance on four domains that reflect hospital quality: the Clinical Process of Care Domain, the Patient Experience of Care domain, the Outcome domain, and the Efficiency domain. The Total Performance Score (TPS) is comprised of the Clinical Process of Care domain score (weighted as 20% of the TPS), the Patient Experience of Care domain score (weighted as 30% of the TPS), the Outcome domain score (weighted as 30% of the TPS), and the Efficiency domain score (weighted as 20% of the TPS).

Below are all Florida Hospitals scored using this 4-measure, weighted system.

```{r totalPerformance, echo=FALSE, warning=FALSE}
ggplot(totalPerformanceScore, aes(x=Hospital.Name, y=domain.score)) + 
  geom_bar(binwidth=1,aes(fill=domain.type),stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_discrete(name = "Domain", labels = c("Experience of Care", "Clinical Process of Care", "Outcome", "Efficiency")) + 
  xlab("Hospital Name") + ylab("Domain Score") + 
  ggtitle("Florida Hospitals Total Performance Score (Weighted)")
```


